```{r project_setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```

# (PART) Project {-}

# Final Project {-}

## Requirements {-}

You will be analyzing a dataset using a regression and a classification analysis. An unsupervised learning analysis is not required for the project, but could be useful. 

**Collaboration:**

- You may work in teams of up to 4 members. Individual work is fine.

<br><br>

**Final deliverables:**

Only one team member has to submit these materials to Moodle. **The due date is Friday, May 5th at 11:59pm CST.**

- Submit a final knitted HTML file (must knit without errors) and corresponding Rmd file containing code for your analysis
  - For cleanliness, please add this R chunk to the top of your Markdown file:
   
```{r}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, tidy = TRUE)
```
   
   
- Include a < 15 minute video presentation of your project that addresses the items in the Grading Rubric below. 
    - Upload the video itself to Moodle. If it's too large, share a link to a recording on the web or in a shared drive.
    - All team members should have an equal speaking role in the presentation.
    - Please use a visual aid (that is, something for the viewer to look at). I think the easiest way to do this is to record a Zoom meeting, then share your screen with some slides.


<br><br>



## Grading Rubric {-}

- Data context (10 points)
    - Clearly describe what the cases in the final clean dataset represent.
    - Broadly describe the variables used in your analyses.
    - Who collected the data? When, why, and how? Answer as much of this as the available information allows.

- Research questions (10 points)
    - Research question(s) for the regression task make clear the outcome variable and its units.
    - Research question(s) for the classification task make clear the outcome variable and its possible categories.

- Regression - Methods (10 points)
    - Describe the regression models you tried.
    - Describe what you did to evaluate models.
        - Indicate how you estimated quantitative evaluation metrics.
        - Indicate what plots you used to evaluate models.

- Regression - Results (10 points)
    - Display evaluation metrics for your different models.
    - Summarize your *final model* and *justify your model choice* using a combination of the following:
        - Analytically: evaluation metrics, plots, variable importance
        - Interpretability/simplicity
    - Show and interpret some representative examples of residual plots for your final model. Does the model show acceptable results in terms of any systematic trends/biases?
    

- Regression - Summary (10 points)
    - Interpret your final model for important predictors, and provide general interpretations about what you learned.
    - Interpret evaluation metrics in context, with units. Is this an acceptable amount of error?
    - All summarizations should be made in original context of the data.

- Classification - Methods (10 points)
    - Describe the classification models you tried.
    - Describe what you did to evaluate models.
        - Indicate how you estimated classification evaluation metrics.
        
- Classification - Results - Variable Importance (10 points)
    - Display evaluation metrics for your different models.
    - Summarize your *final model* and *justify your model choice* using a combination of the following:
        - Analytically: accuracy metrics, variable importance
        - Interpretability/simplicity
        - Plots: ROC, grouped boxplots with probability threshold, etc.
    - Does the model show acceptable results in terms of accuracy, sensitivity, or specificity?


- Classification - Summary (10 points)
    - Interpret your final model for important predictors, and provide general interpretations about what you learned.
    - Interpret evaluation metrics in context. Is this an acceptable accuracy/sensitivity/specificity, relative to the NIR?
    - All summarizations should be made in original context of the data.

- Code (20 points)
    - Knitted, error-free HTML and corresponding Rmd file submitted
    - Code corresponding to all analyses above is present and correct

